\documentclass{article}

\input{preamble}
\input{letterfont}
\input{macros}

\fancyhead[L]{\tbo{Josh Park, Amy Kang, Diya Singh \\ Prof. Kent Quanrud}}
\fancyhead[C]{\tbo{CS 390ATA \\ Homework 6 (\theexercise)}}
\fancyhead[R]{\tbo{Spring 2025 \\ Page \thepage}}

\begin{document}
\setcounter{section}{13}
\setcounter{exercise}{3}
% >>==========|| 13.4 ||==========<<
\begin{exercise}
Let \( G=(V,E) \) be a directed graph.
We say that a set of vertices is \tit{almost independent} if each \( v\in S \) has at most one neighbor in \( S \).\footnote[5]{Two vertices \( u \tand v \) are neighbors if they are connected by an edge.}
Consider the problem of computing the maximum carcdinality of any almost independent set of vertices.
For this problem, either (a) design and analyze a polynomial time algorithm (the faster the better), or (b) prove that a polynomial time algorithm would imply a polynomial time algorithm for SAT.
\end{exercise}

\begin{solution}
  We claim that the maximum cardinality almost independent set (MCAIS) problem is NP-complete.
  To see this, we present a polynomial time reduction from the maximum cardinality independent set (MCIS) problem, a problem we saw to be NP-complete in class, to MCAIS.

  \begin{note}
    Since that the condition for vertices \( v,w\in V \) to be neighbors only requires that there be a connection between the two, we need not be concerned with the directional property of the edges in \( E \).
    Hence, we will interpret \( G \) as an undirected graph.
  \end{note}

  Consider an arbitrary instance of independent set.
  That is, suppose we have a graph \( G=(V,E) \).
  The MCAIS problem seeks the largest \( S\ss V \) such that \( (v,w)\not\in E \) for all \( v,w\in S \).

  Create an auxiliary graph \( G'=(V',E') \), where \( V'=\bigcup\limits_{v_i\in V}\{v_i,v'_i\} \) and \( E'=\{(v_i,v_i')\mid v_i\in V\}\cup E \).
  Thus, \( V' \) contains duplicates of each vertex in \( V \) and \( E' \) contains each edge in \( E \), as well as an additional edge connecting each pair of vertices \( v_j \) and \( v_j' \).
  We can think of \( G' \) as a bilayered graph, where \( G \) is the top layer and every vertex has a copy of itself `hanging' below.
  Importantly, a `hanging' vertex is only connected to the original vertex.

  \begin{subproof}[Correctness.]
    (MCIS\imp MCAIS)
    Assume we have that a maximum cardinality independent set of \( G \) has cardinality \( n \).
    Notice that a subset \( S' \) of \( V' \) containing \( v'_i \) for every \( i \) is independent over \( G' \) with cardinality \( \order V \), since no two hanging vertices are connected by an edge.
    We know there exists some independent set \( S \) of cardinality \( n \) over \( V \).
    Then \( S'' = S\cup S' \) must be almost independent over \( G' \) with cardinality \( \order V + n \), since each vertex in \( S' \) gains at most one neighbor in the top layer of \( G' \).
    Now assume ad absurdum that we have some \( H\sseq V' \) such that \( \order{H} > \order{S''} \).
    If there exists some hanging vertex \( v'_k\not\in H \), then it must be the case that \( v_k\in H \) with some neighboring vertex \( v_{k+1}\in H \).
    However, this implies that \( v'_{k+1}\not\in H \), so we can always replace \( v_k,v_{k+1} \) with their hanging counterparts.
    Thus WLOG every hanging \( v'_i\in H \).
    So \( \order H = \order V + m \) and by hypothesis \( \order V + m > \order V + n \).
    Then we can remove the hanging vertices from \( H \) to get an independent set over \( G \) with cardinality \( m > n \).
    However, this contradicts that the maximum cardinality of an independent set over \( G \) is \( n \).
    Thus \( S'' \) must be a maximum cardinality almost independent set of \( G' \) with cardinality \( \order V + n \).

    (MCIS \pmi MCAIS)
    Assume we have that a maximum cardinality almost independent set \( S \) of \( G' \) has cardinality \( n \).
    As above, we can say WLOG \( S \) contains every hanging vertex.
    Assume ad absurdum that \( S \) contains adjacent vertices \( v_i,v_{i+1} \in V \).
    But \( v'_i \in S \) whence \( v_i \) has two neighbors, a contradiction.
    Thus no two vertices in \( S \) are adjacent in the top layer of \( G' \), which we know to be isomorphic to \( G \).
    Hence by removing every hanging vertex from \( S \), we can get an independent set \( H \) over \( G \) of cardinality \( \order H = n - \order V \).
    Once again, assume ad absurdum that there exists some independent set \( K \) over \( G \) with cardinality \( m > n-\order V \).
    Then by adding every hanging vertex to the set, we obtain an almost independent set with cardinality \( m+\order V > \order H + \order V = n \).
    However, this contradicts that \( n \) is the maximum cardinality for an almost indepedent set over \( G' \) whence \( H \) must be a maximum cardinality independent set over \( G \).
  \end{subproof}
\end{solution}
\pagebreak

\end{document}
